Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-11 shanghai.tex
Moving to a different computer ...

% !TEX encoding = UTF-8 Unicode
\documentclass[11pt]{amsart}
\title[Geometrization of characters of tori]{Geometrization of characters of tori over non-Archimedean local fields}
\usepackage[british]{babel}
\usepackage{datetime}
\date{\today}
\author{Clifton Cunningham}
\address{University of Calgary}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Pacific Institute for the Mathematical Sciences at the University of Calgary}
\email{roed.math@gmail.com}
\usepackage[utf8]{inputenc}
\renewcommand{\baselinestretch}{1.2}
\usepackage[notcite,color]{showkeys}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage[shortalphabetic]{amsrefs}
\usepackage[alphabetic]{amsrefs}
\renewcommand\MR{\relax}
\usepackage{xypic}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{yfonts}
\newcommand{\mathswab}[1]{\operatorname{\textswab{#1}}}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
\usepackage{manfnt}

@@

\begin{abstract}
  We study the category of character sheaves on smooth commutative
  group schemes $G$ locally of finite type over finite fields
  $\Fq$. Assuming that the geometric component group of $G$
  is finitely-generated, we show that the group of isomorphism classes
  of character sheaves on $G$ is canonically isomorphic to the group
  of characters of $G(\Fq)$. We use this result to exhibit an
  isomorphism between admissible characters of $T(K)$ and isomorphism
  classes of character sheaves on the Greenberg transform of the Néron
  model of $T$, where $K$ is any non-Archimdean local field and $T$ is
  any algebraic torus over $K$.
\end{abstract}

@@

%\input{introduction}
%In his celebrated thesis, John Tate writes 
%\begin{quotation}
%    {\it Concerning the characters of $k^*$ [the
%    multiplicative group of a $\mathfrak{p}$-adic field], the
%    situation is different from that of $k^+$.  First of all, we are
%    interested in continuous multiplicative maps $\alpha \mapsto
%    c(\alpha)$ of $k^*$ into the complex numbers, not only in the
%    bounded ones, and shall call such a map a quasi-character,
%    reserving the word ``character'' for the conventional character of
%    absolute value $1$. Secondly, we shall find no model for the group
%    of quasi-characters, or even for the group of characters, though
%    such a model would be of the utmost importance.}
%  \cite{Tate:thesis}*{\S 2.3}
%\end{quotation}
%\noindent
In this paper we show that the monidal category of character sheaves
on the Greenberg transform of the locally finite type Neron model of
$T$ provides an elegant, canonical model for the group of
(quasi-)characters of $T(K)$, where $K$ is any non-Archimedean local
field and $T$ is any algebraic torus over $K$.

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' here. 
Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,\phi)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

Character sheaves on $G$ form a category, denoted by $\CS(G)$. In
fact, $\CS(G)$ is a rigid symmetric monoidal category, so isomorphism
classes of character sheaves on $G$ form a group, denoted by
$\CSiso{G}$; this is the Picard group of the category of
character sheaves on $G$. Our application of character sheaves to
characters of algebraic tori over non-Archimedean local fields rests
on finding the Picard group of character sheaves on $G$, for all smooth commutative group schemes $G$, locally of finite type over $\Fq$, for which the geometric component group is finitely generated.

@@

The main result of this paper is a canonical isomorphism
\[
\CSiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of character sheaves on
$\GN{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \David{I'll do this when we're closer to being finished.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasi-characters of
%$K^\times$: the monoidal category of character sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

Structure of the paper:

Thanks to:

@@

\section{Notation}

@@

For any group $A$, we will denote by $A^*$ the dual group $\Hom_\text{grp}(A, \EEx)$.

\section{The category of character sheaves}\label{section:category}

The category of character sheaves on $G$, denoted by \cdef{$\CS(G)$},
is defined as follows.  A \cdef{character sheaf} on $G$ is a triple
$\cs{L}\ceq (\gcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.0)}{CS.0} $\gcs{L}$ is a constructible $\ell$-adic sheaf on  (the \'etale site of) $\bG$, locally constant of rank-$1$ on each connected component of $\bG$;
\labitem{(CS.1)}{CS.1} $\mu: \bm^* \gcs{L} \to \LxL$
  is an isomorphism of sheaves on $\bGxG$ such that the following diagram commutes, 
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
%
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  \end{tikzcd}
  \]

%
\labitem{(CS.2)}{CS.2} $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{F\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
\end{enumerate}
Morphisms in $\CS(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,\phi)$ and
  $\cs{L'} = (\gcs{L'},\mu',\phi')$ are character sheaves on $G$ then
  $\Hom_{\CS(G)}(\cs{L},\cs{L'})$ is the set of morphisms $\varphi : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \varphi} \arrow[swap]{d}{F} & \Frob{G}^* \gcs{L'} \dar{F'}
  & \arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \varphi} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\varphi} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\varphi\boxtimes \varphi}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

%\begin{remark} \label{rem:CS-action}
The rule $(\gcs{L},\mu,\phi) \mapsto (\gcs{L},\phi)$ defines a forgetful functor from character sheaves on $G$ to $\ell$-adic Weil sheaves on $G$ \cite{Deligne:Weil2}*{Def.~1.1.10 (i)}.

As remarked after \cite{Deligne:Weil2}*{Def.~1.1.10}, a $\ell$-adic Weil sheaf $(\gcs{L},\phi)$ on $G$ may be interpreted as a constructible $\ell$-adic sheaf $\gcs{L}$ on $\bG$ together with an action of the Weil group $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ on $\bG$. Since we will need this point of view later, we review that interpretation here. Setting 
\[
\phi(\Frob{\Fq}^n) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi)
\]
defines, for every $w\in \Weil{\Fq}$, an isomorphism
\[
\phi(w) : w_G^* \gcs{L}\to \gcs{L},
\]
such that
\[
\phi(uv) = \phi(v) \circ v_G^* \phi(u), \qquad \forall u,v\in \Weil{\Fq}.
\]
Note that $\phi = \phi(\Frob{\Fq})$. 
If we further define $\phi_1(w) \ceq (w_G)_*(\phi(w)^{-1})$ then
\[
\phi_1(w) : (w_G)_* \gcs{L}\to \gcs{L},
\]
is an isomorphism and
\[
\phi_1(uv) = \phi_1(u) \circ (u_G)_* \phi_1(v), \qquad \forall u,v\in \Weil{\Fq},
\]
which means the pair $(\gcs{L},\phi_1)$ is almost an action $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of \cite{SGA4.5}*{Expos\'e XIII,~1.1}. The pair $(\gcs{L},\phi_1)$ fails to fulfill the requirements of \cite{SGA4.5}*{Expos\'e XIII,~1.1} only because $\Weil{\Fq}$ is profinite. This fact goes to the heart of the notion of a Weil sheaf and is one of the essential aspects of our definition of $\CS(G)$. 
%\end{remark}

\section{The Picard group of character sheaves}

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$
with duals given by applying the functor 
\Clifton{Should this be $\sheafHom(\ - \ ,\EE)$?} $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$.\footnote{See {\tt http://ncatlab.org/nlab/show/Picard+group}.}
The main result in this paper is the determination of the Picard group of character sheaves on $G$, when the geometric component group of $G$ is finitely generated.

@@

Let $\cs{L}$ be a character sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$ and
therefore an automorphism $\phi_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EEx$, canonically. The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$
guarantees that the function $\trFrob{\cs{L}} : G(\Fq)\to \EEx$,
defined by $g \mapsto \phi_{\bg}$, is a group homomorphism;
we refer to $\trFrob{\cs{L}}$ as the \cdef{trace of Frobenius of $\cs{L}$}.
Although though no trace is involved in the definition, the notion extends 
to any constructible $\ell$-adic Weil sheaf $(\gcs{L},\phi)$ on $G$, in which context,
$\trFrob{(\gcs{L},\phi)} : G(\Fq)\to \EE$ is defined by $\trFrob{(\gcs{L},\phi)}(g) \ceq \trace \phi_{\bg}$.

It is clear that $\trFrob{\cs{L}\otimes \cs{L'}} = \trFrob{\cs{L}} \cdot \trFrob{\cs{L'}}$
(pointwise multiplication of functions) and that $\trFrob{\cs{L}} = \trFrob{\cs{L'}}$
if $\cs{L} \iso \cs{L'}$ in $\CS(G)$. In this way we obtain a group homomorphism
\[
\cdef{\trFrob{G}} : \CSiso{G} \longrightarrow \Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto \trFrob{\cs{L}},
\]
from the Picard group of character sheaves on $G$ to the character group of $G(\Fq)$.

Our main result on the geometrization of admissible characters of
$p$-adic tori [insert internal reference here] follows from a general
fact established later in this paper: if the geometric component group
of $G$ is finitely-generated then the trace of Frobenius defines an
isomorphism from the Picard group of character sheaves $G$ to the
group of characters of $G(\Fq)$ (here, as throughout the paper, $G$ is
a commutative group scheme, smooth and locally of finite type over $\Fq$).

\section{Pull-back of character sheaves}\label{subsection:pullback}

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,\phi) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \CS(G) \to \CS(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \CSiso{G} \rar{f^*} \arrow[swap]{d}{\trFrob{G}} & \CSiso{H} \dar{t_H} \\
  \Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

\begin{proof}
  Let $\cs{L}$ be a character sheaf on $G$. We start by showing that
  $\bf^*\gcs{L}$ is locally constant on connected components of
  $\bH$. Let $c_j :\bH^j \hookrightarrow \bH$ be any
  connected component and let $i$ be the image of $j$ under the group
  homomorphism $\pi_0(\bG) \to \pi_0(\bH)$ obtained by
  applying the functor $\pi_0$ to $\bf$; see \cite{vdG&M}*{III, 3.28}, for example.
  \Clifton{I haven't found a better reference for this yet!}
  Write $c_i : \bG^i \hookrightarrow \bG$ for
  the inclusion of $\bG^i$ and $\bf^j : \bH^j \to \bG^i$
  for the restriction of $f$ to $\bH^j$.  Since $\cs{L}$
  is a character sheaf, $\gcs{L}$ is locally constant on the connect
  components of $\bG$; thus, there is a finite etale covering
  $a_i : X_i \to \bG^i$ such that $a_i^* (\gcs{L}\vert_{\bG^i})$
  is constant. Let $b_j : Y_j \to \bH^j$ be the
  pull-back of $a_i$ along $\bf^j$; let $g_j : Y_j \to X_i$ be
  the map opposite $\bf^j$. Then $b_j$ is a finite etale covering
  of $\bH^j$ and
  \[
  b_j^* \left( (\bf^* \gcs{L})\vert_{\bH^j}\right)
  = (\bf\circ c_j \circ b_j)^*\gcs{L}
  = (c_i\circ a_i\circ g_j)^*\gcs{L} = g_j^* a_i^* (\gcs{L}\vert_{\bG^i}).
  \]

@@

We say that $(\alpha,T)$ in $C(G)$ is a weak isomorphism if $T$ is an isomorphism of vector spaces. 
Let $C(G)[W^{-1}]$ be the localization of $C(G)$ at weak isomorphisms.
Now, let $C_1(G)$ be the subcategory of $C(G)$ such that $\psi$ is a one-dimensional representation
and let $\CSf(G)$ be the localisation of $C_1(G)$ at weak isomorphisms. 
This is the rigid symmetric monoidal category of \cdef{finite character sheaves} on $G$.

>>>>

% !TEX encoding = UTF-8 Unicode
\documentclass[11pt]{amsart}
\title[Geometrization of characters of tori]{Geometrization of characters of tori over non-Archimedean local fields}
\usepackage[british]{babel}
\usepackage{datetime}
\date{\today}
\author{Clifton Cunningham}
\address{University of Calgary}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Pacific Institute for the Mathematical Sciences at the University of Calgary}
\email{roed.math@gmail.com}
\usepackage[utf8]{inputenc}
\renewcommand{\baselinestretch}{1.2}
\usepackage[notcite,color]{showkeys}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage[shortalphabetic]{amsrefs}
\usepackage[alphabetic]{amsrefs}
\renewcommand\MR{\relax}
\usepackage{xypic}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{yfonts}
\newcommand{\mathswab}[1]{\operatorname{\textswab{#1}}}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
\usepackage{manfnt}
%\usepackage{makeidx}

@@

\begin{abstract}
  We introduce and study the category of geometrized characters of smooth commutative
  group schemes $G$ locally of finite type over finite fields
  $\Fq$. Assuming that the geometric component group of $G$
  is finitely-generated, we show that the group of isomorphism classes
  of geometrized characters of $G$ is canonically isomorphic to the group
  of characters of $G(\Fq)$. We use this result to exhibit an
  isomorphism between admissible characters of $T(K)$ and isomorphism
  classes of geometrized characters of the Greenberg transform of the Néron
  model of $T$, where $K$ is any non-Archimdean local field and $T$ is
  any algebraic torus over $K$.
\end{abstract}

@@

%\input{introduction}
%In his celebrated thesis, John Tate writes 
%\begin{quotation}
%    {\it Concerning the characters of $k^*$ [the
%    multiplicative group of a $\mathfrak{p}$-adic field], the
%    situation is different from that of $k^+$.  First of all, we are
%    interested in continuous multiplicative maps $\alpha \mapsto
%    c(\alpha)$ of $k^*$ into the complex numbers, not only in the
%    bounded ones, and shall call such a map a quasi-character,
%    reserving the word ``character'' for the conventional character of
%    absolute value $1$. Secondly, we shall find no model for the group
%    of quasi-characters, or even for the group of characters, though
%    such a model would be of the utmost importance.}
%  \cite{Tate:thesis}*{\S 2.3}
%\end{quotation}
%\noindent
What is a geometrized character?
Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. A geometrized character of $G$ is a triple $\cs{L}\ceq
(\gcs{L},\mu,\phi)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

Geometrized characters of $G$ form a category, denoted by $\GC(G)$. In
fact, $\GC(G)$ is a rigid symmetric monoidal category, so isomorphism
classes of geometrized characters of $G$ form a group, denoted by
$\GCiso{G}$; this is the Picard group of the category of
geometrized characters of $G$. Our application of geometrized characters to
characters of algebraic tori over non-Archimedean local fields rests
on finding the Picard group of geometrized characters of $G$, 
for all smooth commutative group schemes $G$, locally of finite type over $\Fq$, 
for which the geometric component group is finitely generated.

@@

The main result of this paper is a canonical isomorphism
\[
\GCiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of geometrized characters on
$\GN{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \David{I'll do this when we're closer to being finished.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasi-characters of
%$K^\times$: the monoidal category of geometrized characters of the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

In this paper we show that the monoidal category of geometrized characters
on the Greenberg transform of the locally finite type Neron model of
$T$ provides an elegant, canonical model for the group of
(quasi-)characters of $T(K)$, where $K$ is any non-Archimedean local
field and $T$ is any algebraic torus over $K$.

@@

\section{The category of geometrized characters}\label{sec:category}

@@

The category of geometrized characters of $G$, denoted by \cdef{$\GC(G)$},
is defined as follows.  A \cdef{geometrized character} on $G$ is a triple
$\cs{L}\ceq (\gcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.0)}{CS.0} $\gcs{L}$ is a constructible $\ell$-adic sheaf on  (the \'etale site of) $\bG$, locally constant of rank-$1$ on each connected component of $\bG$;
\labitem{(CS.1)}{CS.1} $\mu: \bm^* \gcs{L} \to \LxL$
  is an isomorphism of sheaves on $\bGxG$ such that the following diagram commutes, 
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
%
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  \end{tikzcd}
  \]

%
\labitem{(CS.2)}{CS.2} $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{F\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
\end{enumerate}
Morphisms in $\GC(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,\phi)$ and
  $\cs{L'} = (\gcs{L'},\mu',\phi')$ are geometrized characters of $G$ then
  $\Hom_{\GC(G)}(\cs{L},\cs{L'})$ is the set of morphisms $\alpha : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \alpha} \arrow[swap]{d}{\phi} & \Frob{G}^* \gcs{L'} \dar{\phi'}
  & %\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} 
  & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \alpha} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\alpha} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\alpha\boxtimes \alpha}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

%\begin{remark} \label{rem:CS-action}
The rule $(\gcs{L},\mu,\phi) \mapsto (\gcs{L},\phi)$ defines a forgetful functor from geometrized characters of $G$ to ($\ell$-adic) Weil sheaves on $G$ \cite{Deligne:Weil2}*{Def.~1.1.10 (i)}.
%
As remarked after \cite{Deligne:Weil2}*{Def.~1.1.10}, a Weil sheaf $(\gcs{L},\phi)$ on $G$ may be interpreted as a constructible $\ell$-adic sheaf $\gcs{L}$ on $\bG$ together with an action of the Weil group $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ on $\bG$. Since we will need this point of view later, we review that interpretation here. Setting 
\[
\varphi(\Frob{\Fq}^n) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi)
\]
defines, for every $w\in \Weil{\Fq}$, an isomorphism
\[
\varphi(w) : w_G^* \gcs{L}\to \gcs{L},
\]
such that
\[
\forall u,v\in \Weil{\Fq},\qquad \varphi(uv) = \varphi(v) \circ v_G^* \varphi(u).
\]
Note that $\phi = \varphi(\Frob{\Fq})$. 
If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then
\[
\varphi_1(w) : (w_G)_* \gcs{L}\to \gcs{L},
\]
is an isomorphism and
\[
 \forall u,v\in \Weil{\Fq},\qquad \varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v),
\]
which means the pair $(\gcs{L},\varphi_1)$ is almost an action $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of \cite{SGA4.5}*{Expos\'e XIII,~1.1}. The pair $(\gcs{L},\varphi_1)$ fails to fulfill the requirements of \cite{SGA4.5}*{Expos\'e XIII,~1.1} only because $\Weil{\Fq}$ is profinite. This fact goes to the heart of the notion of a Weil sheaf and is one of the essential aspects of our definition of $\GC(G)$. 
%\end{remark}

\section{The Picard group of geometrized characters}

%\begin{proposition}\label{tensor}
The category $\GC(G)$ of geometrized characters of $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$
with duals given by applying the functor 
\Clifton{Should this be $\sheafHom(\ - \ ,\EE)$?} $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\GC(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\GC(G)$ puts a group
structure on the set $\GCiso{G}$ of isomorphism classes in
$\GC(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$.\footnote{See {\tt http://ncatlab.org/nlab/show/Picard+group}.}
The main result in this paper is the determination of the Picard group of geometrized characters of $G$, when the geometric component group of $G$ is finitely generated.

@@

Let $\cs{L}$ be a geometrized character of $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$ and
therefore an automorphism $\phi_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EEx$, canonically. The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$
guarantees that the function $\trFrob{\cs{L}} : G(\Fq)\to \EEx$,
defined by $g \mapsto \phi_{\bg}$, is a group homomorphism;
we refer to $\trFrob{\cs{L}}$ as the \cdef{trace of Frobenius of $\cs{L}$}.
Although though no trace is involved in the definition, the notion extends 
to any constructible $\ell$-adic Weil sheaf $(\gcs{L},\phi)$ on $G$, in which context,
$\trFrob{(\gcs{L},\phi)} : G(\Fq)\to \EE$ is defined by $\trFrob{(\gcs{L},\phi)}(g) \ceq \trace \phi_{\bg}$.

It is clear that $\trFrob{\cs{L}\otimes \cs{L'}} = \trFrob{\cs{L}} \cdot \trFrob{\cs{L'}}$
(pointwise multiplication of functions) and that $\trFrob{\cs{L}} = \trFrob{\cs{L'}}$
if $\cs{L} \iso \cs{L'}$ in $\GC(G)$. In this way we obtain a group homomorphism
\[
\cdef{\trFrob{G}} : \GCiso{G} \longrightarrow \Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto \trFrob{\cs{L}},
\]
from the Picard group of geometrized characters of $G$ to the character group of $G(\Fq)$.

Our main result on the geometrization of admissible characters of
$p$-adic tori [insert internal reference here] follows from a general
fact established later in this paper: if the geometric component group
of $G$ is finitely-generated then the trace of Frobenius defines an
isomorphism from the Picard group of geometrized characters $G$ to the
group of characters of $G(\Fq)$ (here, as throughout the paper, $G$ is
a commutative group scheme, smooth and locally of finite type over $\Fq$).

\section{Pull-back of geometrized characters}\label{subsection:pullback}

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,\phi) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \GC(G) \to \GC(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \GCiso{G} \rar{f^*} \arrow[swap]{d}{\trFrob{G}} & \GCiso{H} \dar{t_H} \\
  \Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

\begin{proof}
  Let $\cs{L}$ be a geometrized character of $G$. We start by showing that
  $\bf^*\gcs{L}$ is locally constant on connected components of
  $\bH$. Let $c_j :\bH^j \hookrightarrow \bH$ be any
  connected component and let $i$ be the image of $j$ under the group
  homomorphism $\pi_0(\bG) \to \pi_0(\bH)$ obtained by
  applying the functor $\pi_0$ to $\bf$; see \cite{vdG&M}*{III, 3.28}, for example.
  \Clifton{I haven't found a better reference for this yet!}
  Write $c_i : \bG^i \hookrightarrow \bG$ for
  the inclusion of $\bG^i$ and $\bf^j : \bH^j \to \bG^i$
  for the restriction of $f$ to $\bH^j$.  Since $\cs{L}$
  is a geometrized character, $\gcs{L}$ is locally constant on the connect
  components of $\bG$; thus, there is a finite etale covering
  $a_i : X_i \to \bG^i$ such that $a_i^* (\gcs{L}\vert_{\bG^i})$
  is constant. Let $b_j : Y_j \to \bH^j$ be the
  pull-back of $a_i$ along $\bf^j$; let $g_j : Y_j \to X_i$ be
  the map opposite $\bf^j$. Then $b_j$ is a finite etale covering
  of $\bH^j$ and
  \[
  b_j^* \left( (\bf^* \gcs{L})\vert_{\bH^j}\right)
  = (\bf\circ c_j \circ b_j)^*\gcs{L}
  = (c_i\circ a_i\circ g_j)^*\gcs{L} = g_j^* a_i^* (\gcs{L}\vert_{\bG^i}).
  \]

@@

We say that $(\alpha,T)$ in $C(G)$ is a \cdef{weak isomorphism} if $T$ is an isomorphism of vector spaces. 
Let \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at weak isomorphisms.
Now, let \cdef{$C_1(G)$} be the subcategory of $C(G)$ such that $\psi$ is a one-dimensional representation
and let \cdef{$\GCf(G)$} be the localisation of $C_1(G)$ at weak isomorphisms. 
This is the rigid symmetric monoidal category of \cdef{finite geometrized characters} of $G$.
